---
title: "09: Data Frames"
author: "Al Amin Alif"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

## Introduction to Data Frames and dplyr

In R, a data frame is a two-dimensional data structure, similar to a table in a database. The `dplyr` package provides a powerful and intuitive grammar for data manipulation.

```{r}
# Create a data frame
my_dataframe <- data.frame(
  name = c("Charlie", "David", "Emily"),
  age = c(30, 35, 28),
  city = c("New York", "London", "Paris")
)
print(my_dataframe)
```

### Key dplyr verbs

- `select()`: select columns
- `filter()`: filter rows
- `mutate()`: create new columns
- `arrange()`: sort rows
- `summarise()`: summarise data
- `group_by()`: group data

## Exercises

We will be using the `mtcars` dataset for many of these exercises.

1. Load the `mtcars` dataset.

```{r}
# Your code here
data(mtcars)
head(mtcars)
```

2. Select the `mpg`, `cyl`, and `hp` columns from the `mtcars` dataset.

```{r}
# Your code here
select(mtcars, mpg, cyl, hp)
```

3. Filter the `mtcars` dataset to only include cars with more than 6 cylinders.

```{r}
# Your code here
filter(mtcars, cyl>6)
```

4. Create a new column in the `mtcars` dataset called `hp_per_cyl` which is the horsepower per cylinder.

```{r}
# Your code here
mutate(mtcars, hp_per_cyl = hp / cyl)
```

5. Arrange the `mtcars` dataset by the `mpg` column in descending order.

```{r}
# Your code here
arrange(mtcars, desc(mpg))
```

6. Calculate the average `mpg` for each cylinder group.

```{r}
# Your code here
summarise(
  group_by(mtcars, cyl),
  avg_mpg = mean(mpg)
)
```

7. Create an empty data frame named `empty_df`.

```{r}
# Your code here
empty_df <- data.frame()
print(empty_df)
```

8. Create a data frame named `students_df` from the following vectors:
   - `id = c(1, 2, 3)`
   - `name = c("Alice", "Bob", "Charlie")`
   - `score = c(85, 92, 78)`

```{r}
# Your code here
students_df <- data.frame(
  id = c(1, 2, 3),
  name = c("Alice", "Bob", "Charlie"),
  score = c(85, 92, 78)
)
print(students_df)
```

9. Get the structure and a statistical summary of the `students_df` data frame.

```{r}
# Your code here
str(students_df)
summary(students_df)
```

10. Extract the `name` column from `students_df` using the `$` operator.

```{r}
# Your code here
students_df$name
```

11. Extract the first two rows of `students_df`.

```{r}
# Your code here
students_df[1:2, ]
```

12. Extract the third row and the `name` and `score` columns from `students_df`.

```{r}
# Your code here
students_df[3, c("name", "score")]
```

13. Add a new column `grade` to `students_df` where scores >= 90 are "A", scores >= 80 are "B", and others are "C". (Hint: use `mutate` with `case_when`).

```{r}
# Your code here
students_df <- mutate(
  students_df,
  grade = case_when(
    score >= 90 ~ "A",
    score >= 80 ~ "B",
    TRUE ~ "C"
  )
)
print(students_df)
```

14. Remove the `score` column from `students_df`.

```{r}
# Your code here
students_df <- select(students_df, -score)
print(students_df)
```

15. Rename the `name` column in `students_df` to `student_name`.

```{r}
# Your code here
students_df <- rename(students_df, student_name = name)
print(students_df)
```

16. Create another data frame `courses_df` with `id = c(1, 2, 3)` and `course = c("Math", "Science", "History")`. Merge `students_df` and `courses_df` by `id`.

```{r}
# Your code here
courses_df <- data.frame(
  id = c(1, 2, 3),
  course = c("Math", "Science", "History")
)
merged_df <- left_join(students_df, courses_df, by = "id")
print(merged_df)
```

17. Create two data frames, `df1` and `df2`, with identical columns. Combine them row-wise.

```{r}
# Your code here
df1 <- data.frame(A = 1:2, B = c("X", "Y"))
df2 <- data.frame(A = 3:4, B = c("Z", "W"))
combined_rows_df <- bind_rows(df1, df2)
print(df1)
print(df2)
print(combined_rows_df)
```

18. Create two data frames, `df3` and `df4`, with identical rows. Combine them column-wise.

```{r}
# Your code here
df3 <- data.frame(id = 1:2, C = c(10, 20))
df4 <- data.frame(id = 1:2, D = c("P", "Q"))
combined_cols_df <- bind_cols(df3, df4)
print(df3)
print(df4)
print(combined_cols_df)
```

19. Reshape the `mtcars` dataset from wide to long format, keeping `cyl` and `vs` as id variables and gathering `mpg`, `disp`, `hp` into a `metric` and `value` column.

```{r}
# Your code here
mtcars_long <- pivot_longer(
    mtcars,
    cols = c(mpg, disp, hp),
    names_to = "metric",
    values_to = "value"
  )
print(head(mtcars_long))
```

20. Calculate the mean of the `mpg` column in `mtcars`.

```{r}
# Your code here
mean(mtcars$mpg)
```

21. Calculate the number of rows and columns in `mtcars`.

```{r}
# Your code here
nrow(mtcars)
ncol(mtcars)
```

22. Calculate the column means of `mtcars`.

```{r}
# Your code here
colMeans(mtcars)
```

23. Calculate the row sums of `mtcars`.

```{r}
# Your code here
rowSums(mtcars)
```

## Solutions

[Click here for the solutions](dataframes_answers.Rmd)