---
title: "01: Basic Base R Graphics Plots"
author: "Charlie Lindgren"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to Basic Base R Graphics Plots

This document provides examples of how to create basic plots using R's base graphics system. While `ggplot2` is often preferred for its grammar of graphics, understanding base R plotting functions is fundamental.

### Scatter Plot

The `plot()` function is used to create scatter plots in R. The first argument is the x-axis values, and the second argument is the y-axis values. By default, `plot()` creates a scatter plot with points.

```{r}
x <- c(1, 2, 3, 4, 5)
y <- c(2, 4, 6, 8, 10)
plot(x, y)
```

#### Customizing Scatter Plots

Here's an example of a scatter plot with custom point colors and shapes. The `col` argument specifies the color, `pch` specifies the shape, and `cex` specifies the size of the points.

```{r}
plot(x,
  y,
  col = "red",
  pch = 19, # Solid circle
  cex = 2   # Magnification factor
)
```

You can also customize axis labels, titles, and ranges. `xlab` and `ylab` set axis labels, `main` sets the plot title, and `xlim`/`ylim` define axis ranges. `cex.lab` and `cex.axis` control label and axis text size.

```{r}
plot(
  x,
  y,
  col = "blue",
  pch = 19,
  cex = 2,
  xlab = "X-axis Label",
  ylab = "Y-axis Label",
  main = "Custom Scatter Plot Example",
  xlim = c(0, 6),
  ylim = c(0, 12),
  cex.lab = 1.2, # Magnification for axis labels
  cex.axis = 1.2 # Magnification for axis text
)
# Adding a reference line
abline(h = 7, col = "green", lty = 2) # Horizontal line at y=7
# Adding text
text(x = 3, y = 10, "Important Point", col = "purple")
# Adding a legend
legend("topleft", legend = "Data Points", col = "blue", pch = 19)
```

### Line Plot

To create a line plot, use the `type` argument in the `plot()` function, setting `type = "l"`.

```{r}
x_line <- c(1, 2, 3, 4, 5)
y_line <- c(1, 4, 9, 16, 25) # y = x^2
plot(x_line, y_line, type = "l", main = "Line Plot Example")
```

#### Customizing Line Plots

Customize line color and width using `col` and `lwd` (line width).

```{r}
plot(x_line,
  y_line,
  type = "l",
  col = "darkgreen",
  lwd = 3, # Line width
  main = "Custom Line Plot"
)
# Add a horizontal reference line
abline(h = 15, col = "orange", lty = 3) # Dashed line
```

### Bar Plot

The `barplot()` function creates bar plots. The argument is a vector of values.

```{r}
data_vector_bar <- c(15, 22, 18, 30, 25)
barplot(data_vector_bar, main = "Simple Bar Plot")
```

#### Customizing Bar Plots

You can customize colors (`col`), borders (`border`), and provide custom labels (`names.arg`). Set `horiz = TRUE` for horizontal bars.

```{r}
barplot(
  data_vector_bar,
  col = "skyblue",
  border = "darkblue",
  names.arg = c("Category A", "Category B", "Category C", "Category D", "Category E"),
  main = "Custom Vertical Bar Plot",
  xlab = "Categories",
  ylab = "Values"
)

barplot(
  data_vector_bar,
  col = rainbow(length(data_vector_bar)), # Different colors for each bar
  border = "grey",
  names.arg = c("Product X", "Product Y", "Product Z", "Product W", "Product V"),
  horiz = TRUE, # Horizontal bars
  main = "Custom Horizontal Bar Plot",
  xlab = "Values",
  ylab = "Products"
)
```

### Histogram

The `hist()` function creates histograms, showing the distribution of a numeric vector.

```{r}
# Generate random normal data
set.seed(123) # for reproducibility
random_data <- rnorm(1000, mean = 50, sd = 10)
hist(random_data, main = "Simple Histogram")
```

#### Customizing Histograms

Customize the number of bins (`breaks`), colors (`col`), and borders (`border`).

```{r}
hist(random_data,
  breaks = 30, # More bins
  col = "lightgreen",
  border = "darkgreen",
  main = "Custom Histogram with 30 Bins",
  xlab = "Value Range",
  ylab = "Frequency"
)
```

### Box Plot

The `boxplot()` function creates box plots, summarizing the distribution of a dataset using the five-number summary (minimum, first quartile, median, third quartile, maximum).

```{r}
# Generate random data
set.seed(456) # for reproducibility
data_boxplot_1 <- rnorm(100, mean = 0, sd = 1)
boxplot(data_boxplot_1, main = "Simple Box Plot")
```

#### Customizing Box Plots

Customize colors (`col`), borders (`border`), and labels (`names`). You can also plot multiple box plots for comparison.

```{r}
boxplot(
  data_boxplot_1,
  col = "lightpink",
  border = "red",
  names = "Dataset 1",
  main = "Custom Box Plot"
)

# Multiple box plots
set.seed(789) # for reproducibility
data_boxplot_2 <- rnorm(100, mean = 5, sd = 2)
data_boxplot_3 <- rpois(100, lambda = 3) # Poisson distribution

boxplot(
  data_boxplot_1,
  data_boxplot_2,
  data_boxplot_3,
  col = c("lightpink", "lightblue", "lightgreen"),
  border = "grey",
  names = c("Normal 1", "Normal 2", "Poisson"),
  main = "Comparison of Distributions",
  ylab = "Value"
)
```
