---
title: "02: Basic ggplot2 Plots"
author: "Charlie Lindgren"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Introduction to Basic ggplot2 Plots

This document provides an introduction to creating basic plots using `ggplot2`, a powerful and flexible R package for data visualization based on the grammar of graphics. `ggplot2` allows you to build plots layer by layer, providing fine-grained control over every aspect of your visualization.

### Loading Data

We will use the built-in `mtcars` dataset, which contains information about 32 automobiles.

```{r}
data(mtcars)
head(mtcars)
```

### Simple Scatter Plot

Let's create a scatter plot to visualize the relationship between `mpg` (miles per gallon) and `wt` (weight).

The `ggplot()` function initializes a ggplot object. You specify the dataset and map variables to aesthetics (like x-axis, y-axis) using `aes()`. Then, you add geometric objects (geoms) like `geom_point()` for scatter plots using the `+` operator.

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```

This `ggplot2` approach differs significantly from base R plotting. In `ggplot2`, you declare the data and aesthetic mappings once, and then add layers representing different graphical elements. This layered approach makes `ggplot2` highly flexible and extensible.

Compare this to the equivalent base R plot:
```{r}
plot(mtcars$wt, mtcars$mpg)
```

### Improving the Plot: Labels and Theme

We can enhance the plot by adding a title, clearer axis labels, and applying a theme for a more polished look. `labs()` is used for labels and titles, and `theme_classic()` applies a classic white background theme.

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(
    title = "Scatter Plot of Miles per Gallon vs. Weight",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  ) +
  theme_classic()
```

`ggplot2` offers many built-in themes (e.g., `theme_minimal()`, `theme_bw()`) and extensive customization options.

### Further Exploration

This is just a basic introduction. `ggplot2` allows for much more:
*   **Colors and Shapes:** Mapping variables to colors or shapes of points.
*   **Faceting:** Creating multiple plots based on different categories (e.g., `facet_wrap()`, `facet_grid()`).
*   **Other Geoms:** Creating line plots (`geom_line()`), bar plots (`geom_bar()`), histograms (`geom_histogram()`), box plots (`geom_boxplot()`), etc.
*   **Statistical Transformations:** Adding regression lines (`geom_smooth()`), density plots (`geom_density()`).
*   **Customization:** Modifying fonts, colors, legend positions, and more.
