---
title: "04: Advanced ggplot2 Plots"
author: "Charlie Lindgren"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Introduction to Advanced ggplot2 Plots

This document explores some more advanced `ggplot2` functionalities, including adding text labels, using marginal plots with `ggExtra`, and applying custom themes with `ggthemes`.

### Adding Text Labels to Points

Sometimes, it's useful to label individual data points directly on the plot. We can achieve this using `geom_text()`. The `nudge_x` and `nudge_y` arguments help position the labels, and `check_overlap = T` (from `ggrepel` internally, or can be used with `ggrepel::geom_text_repel`) helps prevent text labels from overlapping.

First, let's prepare a subset of the `mtcars` data for cleaner visualization of labels.

```{r}
data_for_labels <- mtcars %>%
  head(10) %>% # Taking the first 10 rows for demonstration
  tibble::rownames_to_column("car_model") # Convert rownames to a column
```

```{r}
ggplot(data_for_labels, aes(x = wt, y = mpg, label = car_model)) +
  geom_point(color = "blue", size = 2) + # Show dots
  geom_text(
    nudge_x = 0.25, # Nudge text slightly to the right
    nudge_y = 0.5,  # Nudge text slightly upwards
    check_overlap = TRUE, # Prevent text from overlapping (basic approach)
    size = 3
  ) +
  labs(
    title = "Miles per Gallon vs. Weight with Car Models",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  ) +
  theme_minimal()
```

### Marginal Plots with `ggExtra`

The `ggExtra` package enhances `ggplot2` plots by adding marginal histograms, density plots, or boxplots to the sides of a scatter plot. This is excellent for visualizing the univariate distributions alongside their bivariate relationship.

First, you might need to install `ggExtra`:
```R
# install.packages("ggExtra")
```

```{r message=FALSE}
library(ggExtra)
```

Let's create a basic scatter plot.

```{r}
p <- ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl), size = hp)) +
  geom_point(alpha = 0.8) +
  labs(
    title = "Scatter Plot with Marginal Distributions",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon",
    color = "Cylinders",
    size = "Horsepower"
  ) +
  theme_bw() +
  theme(legend.position = "bottom") # Move legend for better fit
print(p)
```

Now, let's add different types of marginal plots:

#### Marginal Histogram
```{r}
p_hist <- ggMarginal(
  p, type = "histogram", fill = "skyblue", color = "darkblue"
)
print(p_hist)
```

#### Marginal Density Plot
```{r}
p_density <- ggMarginal(
  p, type = "density", fill = "lightgreen", color = "darkgreen"
)
print(p_density)
```

#### Marginal Box Plot
```{r}
p_boxplot <- ggMarginal(p, type = "boxplot", fill = "salmon", color = "darkred")
print(p_boxplot)
```

### Custom Themes with `ggthemes`

The `ggthemes` package provides a collection of extra themes, geoms, and scales for `ggplot2`, making it easy to apply visually appealing and publication-ready styles. One popular theme is `theme_tufte()`, inspired by Edward Tufte's principles of data visualization.

First, you might need to install `ggthemes`:
```R
# install.packages("ggthemes")
```

```{r message=FALSE}
library(ggthemes)
```

Let's apply the `theme_tufte()` to our previous scatter plot.

```{r}
p_tufte <- ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  labs(
    title = "Miles per Gallon vs. Weight (Tufte Theme)",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon",
    color = "Cylinders"
  ) +
  theme_tufte() + # Apply Tufte theme
  theme(legend.position = "bottom")
print(p_tufte)
```
This demonstrates how easy it is to change the overall aesthetic of your plots using predefined themes.

This concludes the advanced `ggplot2` examples. There's a vast ecosystem of packages that extend `ggplot2`'s capabilities, allowing for highly specialized and customized visualizations.
