---
title: "07: Conditionals"
author: "Al Amin Alif"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to Conditionals

In R, conditionals are used to control the flow of execution based on certain conditions. The most common conditional statement is the `if-else` statement.

```{r}
# An if-else statement
x <- 10
if (x > 5) {
  print("x is greater than 5")
} else {
  print("x is not greater than 5")
}
```

## Exercises

1. Write an `if-else` statement that checks if a number is positive or negative.

```{r}
# Your code here
num <- -5
if(num>=0){
  print("Positive")
}else{
 print("Negative") 
}
```

2. Write an `if-else` statement that checks if a person is eligible to vote (age >= 18).

```{r}
# Your code here
age <- 17
if(age>=18){
  print("Eligible")
}else{
  print("Not Eligible")
}
```

3. Write a function named `check_grade` that takes a numeric grade as input and returns "Pass" if the grade is 50 or above, and "Fail" otherwise.

```{r}
# Your code here
check_grade <- function(grade){
  if(grade>=50){
    print("Pass")
  }else{
    print("Fail")
  }
}
check_grade(47)
```

4. Write a program that checks if a given number is even or odd.

```{r}
# Your code here
odd_even_checker <- function(num){
 if(num%%2 == 0){
   print("Even")
 }else{
   print("Odd")
 }
}
odd_even_checker(3)
```

5. Write a program that checks if a given number is positive, negative, or zero.

```{r}
# Your code here
pos_neg_zero_checker <- function(num){
  if(num == 0){
    print("The number is zero")
  }else if(num>0){
    print("The number is positive")
  }else{
    print("The number is negative")
  }
}
pos_neg_zero_checker(0)
pos_neg_zero_checker(3)
pos_neg_zero_checker(-4)
```

6. Write a program that checks if a given string is empty or not.

```{r}
# Your code here
is_null <- function(str){
  if(nchar(str)==0){
    print("String is empty")
  }else{
    print("String is not empty")
  }
}
is_null("")
is_null("string")
```

7. Write a program that checks if a given vector contains a specified element.

```{r}
# Your code here
input_vec <- c(1,2,3,4,5)
element_checker <- function(elem){
  if(elem %in% input_vec){
    print("Vector contains the element")
  }else{
    print("Vector does not contain the element")
  }
}
element_checker(6)
element_checker(3)
```

8. Write a program that checks if a given number is between two other numbers.

```{r}
# Your code here
is_between <- function(num, lower, upper){
  if(num>lower && num<upper){
    print("Is between")
  }else{
    print("Is NOT between")
  }
}
is_between(3,4,5)
is_between(3,1,5)
```

9. Write a program that checks if a given year is a leap year.

```{r}
# Your code here
is_leap_year <- function(year){
  if((year %% 4 == 0 && year %% 100 != 0) || (year %% 400 == 0)){
    print("Leap year")
  }else{
    print("Not leap year")
  }
}
is_leap_year(2025)
is_leap_year(2024)
```

10. Write a program that checks if a given number is a prime number.

```{r}
# Your code here
is_prime <- function(num) {
  if (num <= 1) {
    print("Not a prime number")
    return()
  }
  for (i in 2:(num - 1)) {
    if (num %% i == 0) {
      print("Not a prime number")
      return()
    }
  }
  print("Prime number")
}
is_prime(7)
is_prime(10)
```

11. Write a program that assigns grades (A, B, C, D, F) based on a given score (0-100).

```{r}
# Your code here
assign_grade <- function(score) {
  if (score >= 90) {
    print("A")
  } else if (score >= 80) {
    print("B")
  } else if (score >= 70) {
    print("C")
  } else if (score >= 60) {
    print("D")
  } else {
    print("F")
  }
}
assign_grade(95)
assign_grade(72)
assign_grade(42)
```

12. Write a program that checks if a given number is a perfect square.

```{r}
# Your code here
is_perfect_square <- function(num){
  if(num<0){
    print("Not perfect square")
    return()
  }
  sqrt_num <- sqrt(num)
  if(sqrt_num == floor(sqrt_num)){
    print("Perfect square")
  }else{
    print("Not perfect square")
  }
}
is_perfect_square(16)
is_perfect_square(12)
```

13. Write a program that checks if a given number is a palindrome.

```{r}
# Your code here
is_palindrome <- function(num) {
  num_str <- as.character(num)
  reversed_str <- paste(rev(strsplit(num_str, "")[[1]]), collapse = "")
  if (num_str == reversed_str) {
    print("Palindrome")
  } else {
    print("Not a Palindrome")
  }
}
is_palindrome(121)
is_palindrome(123)
```

## Solutions

[Click here for the solutions](conditionals_answers.Rmd)