---
title: "Analyzing Cosmetics Data (CSV)"
author: "Charlie Lindgren"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
```

## Introduction

This document demonstrates how to load, explore, transform, and visualize cosmetics product data from a CSV file using R, with a focus on `dplyr` and `ggplot2` for modern data analysis workflows.

## Exercises

### Exercise 1: Load the Data

Load the `most_used_beauty_cosmetics_products_extended.csv` dataset into an R data frame. Use `read_csv()` from the `readr` package for this.

```{r}
# Your code here
```

### Exercise 2: Explore the Data

Display the structure of the `cosmetics` data frame, get a statistical summary, and check the unique values of the `Product_Size` column.

```{r}
# Your code here
```

### Exercise 3: Data Transformation - Product Size

Transform the `Product_Size` column into an ordered factor with appropriate levels.

```{r}
# Your code here
```

### Exercise 4: Data Filtering

Retrieve all products that have a `Product_Size` in excess of "150ml". Store this in a new data frame called `cosmetics_large_size`.

```{r}
# Your code here
```

### Exercise 5: Data Summarization

Get a summary of the `cosmetics_large_size` data frame and find out how many rows it contains.

```{r}
# Your code here
```

### Exercise 6: Data Visualization - Price Distribution

Create a histogram of the `Price_USD` column to visualize its distribution.

```{r}
# Your code here
```

## Solutions

[Click here for the solutions](analyze_cosmetics_answers.Rmd)
