---
title: "R Programming Assignment: Comprehensive Data Analysis"
author: "Your Name Here"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required packages
if (!require("testthat")) install.packages("testthat")
library(testthat)

# Load the dataset
data <- read.csv("student_scores.csv", stringsAsFactors = FALSE)
```

# Introduction

Welcome to the R Programming Assignment! This notebook contains **10 functions** you need to implement.

**How this works:**

1. Each function has a **stub** (empty function body) that you need to complete
2. After each function, there are **PUBLIC TESTS** that give you immediate feedback
3. Run the tests frequently to check your progress
4. Your instructor has additional **HIDDEN TESTS** that check edge cases and logic

**Grading:**

- Public tests help you verify basic functionality
- Hidden tests (run by instructor) check complete correctness
- You must pass ALL hidden tests to get full credit

**Important:** Functions must handle edge cases properly. Read requirements carefully!

---

# Part 1: Data Loading & Validation

## Function 1: Load and Validate Student Data

**Task:** Load a CSV file and validate it has the expected structure.

**Requirements:**

- Return NULL if file doesn't exist (don't throw error)
- Return NULL if file is empty
- Return NULL if required columns are missing: "student_id", "name", "math_score"
- Return the data frame only if all validations pass
- Convert enrollment_date to Date type if present

```{r load_student_data}
# YOUR CODE HERE
load_student_data <- function(filename) {
  # TODO: Implement with proper validation
  NULL
}
```

### Public Tests - Function 1

```{r test_load_student_data, error=TRUE}
public_test_passed <- tryCatch({
  test_that("load_student_data works", {
    result <- load_student_data("student_scores.csv")
    expect_false(is.null(result), info = "Function returned NULL - did you implement it?")
    expect_true(is.data.frame(result))
    expect_true("student_id" %in% colnames(result))
    expect_true("name" %in% colnames(result))
    expect_true("math_score" %in% colnames(result))
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for load_student_data!\n")
} else {
  cat("✗ Public tests FAILED for load_student_data. Check the error messages above.\n")
}
```

---

## Function 2: Filter Students with Incomplete Data

**Task:** Return students who have missing values (NA) in ANY of the score columns.

**Requirements:**

- Check math_score, english_score, science_score columns
- Return data frame with ALL original columns
- Include a new column "missing_count" showing how many scores are missing (1, 2, or 3)
- Sort by missing_count (descending), then by student_id (ascending)
- If no students have missing data, return empty data frame with correct column structure

```{r filter_incomplete_students}
# YOUR CODE HERE
filter_incomplete_students <- function(data) {
  # TODO: Find students with ANY missing scores
  NULL
}
```

### Public Tests - Function 2

```{r test_filter_incomplete_students, error=TRUE}
public_test_passed <- tryCatch({
  test_that("filter_incomplete_students works", {
    result <- filter_incomplete_students(data)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    expect_true("missing_count" %in% colnames(result))
    # Test data has students with missing scores, so result should not be empty
    expect_true(nrow(result) > 0, info = "Should find students with missing scores")
    expect_true(all(result$missing_count >= 1 & result$missing_count <= 3))
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for filter_incomplete_students!\n")
} else {
  cat("✗ Public tests FAILED for filter_incomplete_students.\n")
}
```

---

# Part 2: Statistical Analysis with Edge Cases

## Function 3: Calculate Weighted Average Score

**Task:** Calculate a weighted average of the three exam scores.

**Requirements:**

- Weights: math (40%), english (30%), science (30%)
- If a student has NA in a score, redistribute that weight proportionally to other scores
  - Example: If science is NA, use math (57.14%) and english (42.86%)
  - If two scores are NA, use the one available score (100%)
- If all three scores are NA, return NA
- Return a numeric vector with one value per student
- Round to 2 decimal places

```{r calculate_weighted_average}
# YOUR CODE HERE
calculate_weighted_average <- function(data) {
  # TODO: Calculate weighted average with weight redistribution
  NULL
}
```

### Public Tests - Function 3

```{r test_calculate_weighted_average, error=TRUE}
public_test_passed <- tryCatch({
  test_that("calculate_weighted_average works", {
    result <- calculate_weighted_average(data)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_type(result, "double")
    expect_equal(length(result), nrow(data))
    expect_true(sum(!is.na(result)) > 0)
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for calculate_weighted_average!\n")
} else {
  cat("✗ Public tests FAILED for calculate_weighted_average.\n")
}
```

---

## Function 4: Identify Score Anomalies

**Task:** Find students whose performance varies significantly across subjects.

**Requirements:**

- Calculate the range (max - min) of each student's three exam scores
- Return students where range > threshold AND they have scores in at least 2 subjects
- Add column "score_range" with the calculated range
- Add column "highest_subject" (one of: "math_score", "english_score", "science_score")
- Add column "lowest_subject" (one of: "math_score", "english_score", "science_score")
- If student has all NA scores, exclude them
- Sort by score_range (descending)

```{r identify_score_anomalies}
# YOUR CODE HERE
identify_score_anomalies <- function(data, threshold) {
  # TODO: Find students with high score variance
  NULL
}
```

### Public Tests - Function 4

```{r test_identify_score_anomalies, error=TRUE}
public_test_passed <- tryCatch({
  test_that("identify_score_anomalies works", {
    result <- identify_score_anomalies(data, 15)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    expect_true("score_range" %in% colnames(result))
    expect_true("highest_subject" %in% colnames(result))
    expect_true("lowest_subject" %in% colnames(result))
    if (nrow(result) > 0) {
      expect_true(all(result$score_range > 15))
    }
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for identify_score_anomalies!\n")
} else {
  cat("✗ Public tests FAILED for identify_score_anomalies.\n")
}
```

---

# Part 3: String Manipulation & Data Cleaning

## Function 5: Standardize Student Names

**Task:** Clean and standardize the name format.

**Requirements:**

- Convert to "LastName, FirstName" format
- Handle names with titles (Dr., Mr., Ms., etc.) - remove them
- Handle middle names - include them in FirstName part
- Trim extra whitespace
- Capitalize properly (Title Case)
- If name has only one word, return "Unknown, [Name]"
- Return the original data frame with "name" column updated

**Examples:**

- "Alice Johnson" → "Johnson, Alice"
- "Dr. Bob Smith" → "Smith, Bob"
- "Mary Jane Watson" → "Watson, Mary Jane"
- "Prince" → "Unknown, Prince"

```{r standardize_names}
# YOUR CODE HERE
standardize_names <- function(data) {
  # TODO: Reformat names to LastName, FirstName
  NULL
}
```

### Public Tests - Function 5

```{r test_standardize_names, error=TRUE}
public_test_passed <- tryCatch({
  test_that("standardize_names works", {
    result <- standardize_names(data)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    expect_true("name" %in% colnames(result))
    expect_true(any(grepl(",", result$name)))
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for standardize_names!\n")
} else {
  cat("✗ Public tests FAILED for standardize_names.\n")
}
```

---

## Function 6: Extract and Categorize Email Domains

**Task:** Create email addresses and categorize by domain type.

**Requirements:**

- Generate email: firstname.lastname@domain.edu (all lowercase, no spaces)
- Remove any special characters from names for email
- Determine domain based on major:
  - "Computer Science", "Engineering" → "tech.university.edu"
  - "Mathematics", "Physics" → "science.university.edu"
  - "Biology", "Chemistry" → "bio.university.edu"
  - Others → "general.university.edu"
- Add column "email" with generated email
- Add column "domain_category" (one of: "tech", "science", "bio", "general")
- Return the data frame with new columns

```{r create_email_addresses}
# YOUR CODE HERE
create_email_addresses <- function(data) {
  # TODO: Generate emails based on name and major
  NULL
}
```

### Public Tests - Function 6

```{r test_create_email_addresses, error=TRUE}
public_test_passed <- tryCatch({
  test_that("create_email_addresses works", {
    result <- create_email_addresses(data)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    expect_true("email" %in% colnames(result))
    expect_true("domain_category" %in% colnames(result))
    expect_true(all(grepl("@.*\\.edu$", result$email)))
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for create_email_addresses!\n")
} else {
  cat("✗ Public tests FAILED for create_email_addresses.\n")
}
```

---

# Part 4: Complex Conditional Logic

## Function 7: Assign Letter Grades with Curve

**Task:** Assign letter grades using a curved grading system.

**Requirements:**

- Calculate each student's average score (mean of available exam scores)
- Apply curve: New_avg = min(100, avg + curve_points)
- Assign grades based on curved average:
  - A: >= 90
  - B: 80-89
  - C: 70-79
  - D: 60-69
  - F: < 60
- If student has all NA scores, assign "I" (Incomplete)
- Return character vector of grades
- Grades must be in same order as input data

```{r assign_curved_grades}
# YOUR CODE HERE
assign_curved_grades <- function(data, curve_points) {
  # TODO: Calculate curved grades
  NULL
}
```

### Public Tests - Function 7

```{r test_assign_curved_grades, error=TRUE}
public_test_passed <- tryCatch({
  test_that("assign_curved_grades works", {
    result <- assign_curved_grades(data, 5)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_type(result, "character")
    expect_equal(length(result), nrow(data))
    valid_grades <- c("A", "B", "C", "D", "F", "I")
    expect_true(all(result %in% valid_grades))
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for assign_curved_grades!\n")
} else {
  cat("✗ Public tests FAILED for assign_curved_grades.\n")
}
```

---

## Function 8: Predict At-Risk with Multiple Criteria

**Task:** Identify at-risk students using multiple weighted criteria.

**Requirements:**

- Calculate risk score (0-100) based on:
  - Low attendance (< 80%): +30 points
  - Low average exam score (< 70): +25 points
  - Failed midterm (passed_midterm == FALSE): +20 points
  - Missing assignments (any NA in assignment scores): +15 points
  - High score variance (range > 20): +10 points
- Students with risk_score >= risk_threshold are "at risk"
- Return data frame with at-risk students only
- Add columns: "risk_score" and "risk_factors" (comma-separated list of triggered factors)
- Sort by risk_score (descending)

**Example risk_factors:** "Low Attendance, Failed Midterm, Missing Assignments"

```{r predict_at_risk_students}
# YOUR CODE HERE
predict_at_risk_students <- function(data, risk_threshold) {
  # TODO: Calculate risk scores with multiple criteria
  NULL
}
```

### Public Tests - Function 8

```{r test_predict_at_risk_students, error=TRUE}
public_test_passed <- tryCatch({
  test_that("predict_at_risk_students works", {
    result <- predict_at_risk_students(data, 40)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    expect_true("risk_score" %in% colnames(result))
    expect_true("risk_factors" %in% colnames(result))
    if (nrow(result) > 0) {
      expect_true(all(result$risk_score >= 40))
    }
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for predict_at_risk_students!\n")
} else {
  cat("✗ Public tests FAILED for predict_at_risk_students.\n")
}
```

---

# Part 5: Advanced Data Aggregation

## Function 9: Create Detailed Major Statistics

**Task:** Generate comprehensive statistics by major with conditional calculations.

**Requirements:**

- Group by major
- Calculate for each major:
  - `student_count`: Total students
  - `avg_attendance`: Mean attendance
  - `pass_rate`: Percentage with average score >= 60
  - `honor_roll_count`: Number with average score >= 85
  - `perfect_attendance`: Number with attendance >= 95%
  - `completion_rate`: Percentage who completed all assignments (no NAs)
  - `avg_assignments`: Mean of all assignment scores
- Round all percentages to 1 decimal place
- Round all averages to 2 decimal places
- Sort by pass_rate (descending), then by major (alphabetical)

```{r create_major_statistics}
# YOUR CODE HERE
create_major_statistics <- function(data) {
  # TODO: Calculate comprehensive statistics by major
  NULL
}
```

### Public Tests - Function 9

```{r test_create_major_statistics, error=TRUE}
public_test_passed <- tryCatch({
  test_that("create_major_statistics works", {
    result <- create_major_statistics(data)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    required_cols <- c("major", "student_count", "avg_attendance", "pass_rate",
                       "honor_roll_count", "perfect_attendance", "completion_rate")
    expect_true(all(required_cols %in% colnames(result)))
    expect_true(nrow(result) > 0)
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for create_major_statistics!\n")
} else {
  cat("✗ Public tests FAILED for create_major_statistics.\n")
}
```

---

## Function 10: Generate Performance Report with Ranking

**Task:** Create a ranked performance report with percentile calculations.

**Requirements:**

- Calculate overall_score for each student: weighted average of exam scores (math: 40%, english: 30%, science: 30%)
- Add columns:
  - `overall_score`: The calculated weighted average
  - `class_rank`: Rank based on overall_score (1 = highest, ties get same rank)
  - `percentile`: Percentile rank (0-100, round to 1 decimal)
  - `performance_tier`: Based on percentile:
    - "Top 10%": percentile >= 90
    - "Top 25%": percentile >= 75
    - "Top 50%": percentile >= 50
    - "Below Average": percentile < 50
  - `improvement_needed`: TRUE if attendance < 85% OR any exam score < 65
- Exclude students with all NA exam scores
- Sort by class_rank (ascending)
- Return data frame with: student_id, name, overall_score, class_rank, percentile, performance_tier, improvement_needed

```{r generate_performance_report}
# YOUR CODE HERE
generate_performance_report <- function(data) {
  # TODO: Create ranked performance report
  NULL
}
```

### Public Tests - Function 10

```{r test_generate_performance_report, error=TRUE}
public_test_passed <- tryCatch({
  test_that("generate_performance_report works", {
    result <- generate_performance_report(data)
    expect_false(is.null(result), info = "Function returned NULL")
    expect_true(is.data.frame(result))
    required_cols <- c("student_id", "name", "overall_score", "class_rank",
                       "percentile", "performance_tier", "improvement_needed")
    expect_true(all(required_cols %in% colnames(result)))
    expect_true(nrow(result) > 0)
    expect_true(all(result$percentile >= 0 & result$percentile <= 100))
  })
  TRUE
}, error = function(e) FALSE)

if (public_test_passed) {
  cat("✓ Public tests passed for generate_performance_report!\n")
} else {
  cat("✗ Public tests FAILED for generate_performance_report.\n")
}
```

---

# Summary

Run all chunks above to test your implementations. Make sure all public tests pass!

**Before submitting:**

1. Knit this document to HTML to ensure everything runs
2. Check that you see "✓ Public tests passed" for all 10 functions
3. Submit this .Rmd file along with your AI usage declaration

**Note:** Your instructor will run additional hidden tests that check:

- Edge cases (empty data, all NAs, boundary values)
- Correct logic implementation
- Proper handling of special cases
- Weight redistribution accuracy
- Ranking and percentile calculations
- String manipulation correctness

Good luck!