---
title: "01: Vectors"
author: "Al Amin Alif"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to Vectors

In R, a vector is a sequence of data elements of the same basic type. The `c()` function is used to create a vector.

```{r}
# A numeric vector
numeric_vector <- c(1, 2, 3, 4, 5)
print(numeric_vector)

# A character vector
character_vector <- c("a", "b", "c", "d", "e")
print(character_vector)
```

## Exercises

1. Create a numeric vector named `my_numeric_vector` with the numbers 10, 20, 30, 40, and 50.

```{r}
# Your code here
my_numeric_vector <- c(10, 20, 30, 40, 50)
print(my_numeric_vector)
```

2. Create a character vector named `my_character_vector` with the names of your three favorite colors.

```{r}
# Your code here
my_character_vector <- c("blue", "green", "yellow")
print(my_character_vector)
```

3. Get the length of `my_numeric_vector`.

```{r}
# Your code here
length(my_numeric_vector)
```

4. Create a vector containing a numeric, a logical, and a character element. What is the type of the resulting vector?

```{r}
# Your code here
mixed_character_vector <- c(10, "Leo", TRUE)
typeof(mixed_character_vector)
print(mixed_character_vector)
```

5. Add the following two vectors: `c(1, 2, 3)` and `c(4, 5, 6)`.

```{r}
# Your code here
added_vectors <- c(1, 2, 3) + c(4, 5, 6)
print((added_vectors))
```

6. What happens if you add two vectors of different lengths? Try adding `c(1, 2, 3)` and `c(4, 5)`.

```{r}
# Your code here
mismatched_vectors_added <- c(1, 2, 3) + c(4, 5)
print(mismatched_vectors_added)
```

7. Create an empty vector named `my_results`. Then, append the number 10 to it.

```{r}
# Your code here
my_results <- c()
my_results <- c(my_results, 10)
print(my_results)
```

8. Given the vector `c(1, 5, 2, 8, 4, NA)`, find the sum, mean, and product of the vector. Then, do the same but ignore the `NA` value.

```{r}
# Your code here
my_vector <- c(1, 5, 2, 8, 4, NA)
sum(my_vector)
mean(my_vector)
prod(my_vector)
sum(my_vector, na.rm = TRUE)
mean(my_vector, na.rm = TRUE)
prod(my_vector, na.rm = TRUE)
```

9. Given the vector `c(1, 5, 2, 8, 4)`, find the minimum and maximum values.

```{r}
# Your code here
my_vector2 <- c(1, 5, 2, 8, 4)
min(my_vector2)
max(my_vector2)
```

10. Sort the vector `c(1, 5, 2, 8, 4)` in ascending and descending order.

```{r}
# Your code here
my_vector3 <- c(1, 5, 2, 8, 4)
sort(my_vector3)
sort(my_vector3, decreasing = TRUE)
```

11. Check if the number 5 is present in the vector `c(1, 5, 2, 8, 4)`.

```{r}
# Your code here
5 %in% c(1, 5, 2, 8, 4)
6 %in% c(1, 5, 2, 8, 4)
```

## Solutions

[Click here for the solutions](vectors_answers.Rmd)