---
title: "Example R Markdown: Student Grade Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This notebook demonstrates basic data analysis in an **R Markdown** file.
Compare this with `example_r_script.R` to see the difference!

**How to run:**

- Single chunk: Click ▶ or `Cmd+Shift+Enter` (Mac) / `Ctrl+Shift+Enter` (Windows)
- Single line: `Cmd+Enter` (Mac) / `Ctrl+Enter` (Windows)
- Knit to HTML: Click "Knit" or `Cmd+Shift+K` (Mac) / `Ctrl+Shift+K` (Windows)

---

# 1. Setup

First, we create our sample data. In R Markdown, code lives inside "chunks":

```{r create-data}
# Create sample data (normally you'd load from a file)
students <- data.frame(
  name = c("Alice", "Bob", "Carol", "David", "Eve"),
  math = c(92, 78, 85, 91, 88),
  english = c(88, 82, 90, 76, 94),
  science = c(95, 75, 82, 88, 91)
)

# Display the data (in R Markdown, this renders as a nice table)
students
```

Notice how the output appears right below the code chunk!

---

# 2. Calculate Averages

Let's add an average column for each student:

```{r calculate-averages}
# Calculate each student's average across subjects
students$average <- rowMeans(students[, c("math", "english", "science")])

# Round to 1 decimal place
students$average <- round(students$average, 1)

students
```

The class average is **`r round(mean(students$average), 1)`** — notice how we can embed R code directly in the text using backticks!

---

# 3. Assign Letter Grades

Now we'll write a function to assign letter grades:

```{r grade-function}
# Function to convert score to letter grade
assign_grade <- function(score) {
  if (score >= 90) return("A")
  if (score >= 80) return("B")
  if (score >= 70) return("C")
  if (score >= 60) return("D")
  return("F")
}
```

Apply it to each student:

```{r apply-grades}
students$grade <- sapply(students$average, assign_grade)
students
```

---

# 4. Summary Statistics

Let's look at some summary statistics:

```{r summary-stats}
# Class average
class_avg <- mean(students$average)

# Top and bottom students
top_student <- students$name[which.max(students$average)]
bottom_student <- students$name[which.min(students$average)]
```

**Results:**

| Metric | Value |
|--------|-------|
| Class Average | `r round(class_avg, 1)` |
| Top Student | `r top_student` |
| Needs Improvement | `r bottom_student` |

**Grade Distribution:**

```{r grade-distribution}
table(students$grade)
```

---

# 5. Visualization

R Markdown displays plots directly in the document:

```{r plot, fig.width=8, fig.height=5}
# Bar plot of averages
barplot(
  students$average,
  names.arg = students$name,
  main = "Student Averages",
  ylab = "Average Score",
  col = "steelblue",
  ylim = c(0, 100)
)

# Add a line for class average
abline(h = class_avg, col = "red", lty = 2, lwd = 2)
legend("topright", "Class Average", col = "red", lty = 2, lwd = 2)
```

---

# Conclusion

We analyzed `r nrow(students)` students and found:

- **`r sum(students$grade == "A")`** students earned an A
- **`r sum(students$grade == "B")`** students earned a B
- The class average was **`r round(class_avg, 1)`**

---

# Notes on .Rmd Files

## Pros

- **Rich formatting** — Headers, bold, tables, lists
- **Code + output together** — Results appear right after code
- **Publishable documents** — Knit to HTML, PDF, or Word
- **Inline R code** — Embed calculations in text
- **Reproducible** — Knitting runs everything fresh

## Cons

- More complex syntax than plain .R
- Knitting can be slow for large analyses
- Need to understand Markdown formatting
- Debugging can be trickier (chunk ordering matters)

## When to use .Rmd

- Reports and assignments
- Sharing analysis with others
- Documentation with code examples
- When you want formatted output
- Reproducible research
